### 2013-10-20 src2/thesis.tex
small update

Paper E, was the first I wrote with Josef Svenningsson who had been thinking about an interesting
variation of counting sort and wanted to implement it. Together we added atomic operations to 
the deep embedding and I modified the Obsidian code generator to accomodate them. Sections 
1 and 2 in the paper are written by Josef. Sections 3 through 7 are mostly my work, including 
the benchmarking. Section 8 is joint work.
%Me and Josef are equal partners in this paper and got much valuable advice from 
%Mary Sheeran.

@@

Modern CPUs are parallel in a number of different ways. They discover instruction
level parallelism (ILP), they have multiple processor cores and vector units for 
Single Instruction Multiple Data (SIMD)  parallelism. The exact configuration, number
of cores and number of vector units, varies over available processor models. 
To obtain maximum performance, with a specific target processor in mind, the programmer
can use low-level intrinsics in order to get SIMD parallelism and a threading library
for parallelism over the available cores. One problem with this is that the resulting program 
would be specialised to a specific processor model. The Intel Array Building Blocks (ArBB) 
system approached this problem by providing a high level programming interface~\citet{ARBB2011}.
This interface is based on a set of parallel operations on arrays, such as map, folds 
and scans. ArBB was implemented as an embedded language in C++ but a low-level C interface to 
its capabilities was also offered. ArBB programs are JIT-ed (Just-In-Time compiled) to 
target the available CPU or accelerator (Such as the Intel Larrabee). In 
chapter~\ref{chap:ArBB}, there are papers describing work towards embedding Intel 
ArBB in Haskell.

@@

With EmbArBB we improve the interfacing with Haskell by implementing a deeply embedded language. 
Now, programs in EmbArBB create ASTs. These ASTs are compiled via the ArBB bindings creating 
function objects that be run.

\noindent\emph{Short section about the approach taken with EmbArBB}

@@

%\subsection{The results}

>>>>

Paper E, was the first I wrote with Josef Svenningsson who had been thinking about an interesting
variation of counting sort and wanted to implement it. Together we added atomic operations to 
Obsidian and I modified the code generator to accomodate them. Sections 
1 and 2 in the paper are written by Josef. Sections 3 through 7 are mostly my work, including 
the benchmarking. Section 8 is joint work.
%Me and Josef are equal partners in this paper and got much valuable advice from 
%Mary Sheeran.

@@

While it is sometimes possible to automatically parallelise programs originally 
written for serial execution, it is becoming clear that explicitly parallel 
programming models are needed for making use of all parallel resources~\cite{STRUCTURED}. 
This also means that it is becoming important to think in parallel from the beginning.

Modern CPUs are parallel in a number of different ways. They discover instruction
level parallelism (ILP), they have multiple processor cores and vector units for 
Single Instruction Multiple Data (SIMD)  parallelism. The exact configuration, number
of cores and number of vector units, varies over available processor models. 
To obtain maximum performance, with a specific target processor in mind, the programmer
can use low-level intrinsics in order to get SIMD parallelism and a threading library
for parallelism over the available cores. One problem with this is that the resulting program 
would be specialised to a specific processor model. The Intel Array Building Blocks (ArBB) 
system approach this problem by providing a high level programming interface~\citet{ARBB2011}.
This interface is based on a set of parallel operations on arrays, such as map, folds 
and scans. ArBB was implemented as an embedded language in C++ but a low-level C interface to 
its capabilities was also offered. ArBB programs are JIT-ed (Just-In-Time compiled) to 
target the available CPU or accelerator (Such as the Intel Larrabee). In 
chapter~\ref{chap:ArBB}, there are papers describing work towards embedding Intel 
ArBB in Haskell.

@@

With EmbArBB we improve the interfacing with Haskell by implementing a deeply embedded language. 
Now, programs in EmbArBB create ASTs. These ASTs are compiled via the ArBB bindings creating 
function objects that can then be run.

\subsubsection{Dense and nested arrays}

ArBB supports one, two and three dimensional arrays, called dense vectors. In EmbArBB we 
represent these with a data type, {\tt DVector}, parameterised by dimensionality and 
element type. The dimension parameter can be either {\tt Dim1}, {\tt Dim2} or {\tt Dim3}. 
There are also nested vectors, in EmbArBB called {\tt NVector}, parameterised on element type.

\subsubsection{Deep embedding}

\begin{verbatim}
data Expr = Lit Literal
          | Var Variable

          | Index0 Expr 
          | ResIndex Expr Int

          | Map (R GenRecord) [Expr]  
          | Call (R GenRecord) [Expr]

          | While ([Expr] -> Expr)  ([Expr] -> [Expr])  [Expr]

          | While2 ([Expr] -> Expr)  
                   ([Expr] -> [Expr])  
                   [Expr}

          | If Expr Expr Expr

          | Op Op [Expr]   
\end{verbatim}

\subsubsection{compilation}

\emph{Sharing detection}

\emph{let ArBB do the optimisations}

\emph{having deep embedding means we could compile to something else}

\subsubsection{Evaluation}

@@
